Some figures to illustrate basic ideas in FDR

### 2016-12-09 analysis/makefig_FDReg.Rmd
knit makefig_FDReg.Rmd (note that this is in RStudio 1.0.44).

```{r chunk-options, include=FALSE}
require(ashr)
require(qvalue)
require(locfdr)
require(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```

@@

Generates DReg_hist.pdf?

@@

# Roughly compute a local fdr for qvalue (to aid the plotting of the
# decomposition in zscore space).
temp     = hist(pval,nclass = 50)
bin_fdr  = res.qvalue$pi0/temp$density # compute local fdr in each bin of histogram
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr) # threshold at 1, so fdr never bigger than 1
qval_fdr = qvalue::lfdr(pval)
```
Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:

```{r plot_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 1,title = "p values",
                cex.axis = 0.8,cex.main = 0.8)

```

Now the decomposition:

@@

@@

# p value histograms
altnullhist(pval,qval_fdr,main="p values",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)

# z score histograms
nullalthist(zscore,qval_fdr,main="z scores",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.locfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="",ncz=50,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
```

@@

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
#pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol=4,byrow=TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz=25

#p value histograms
altnullhist(pval,qval_fdr,main="p values: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8)
#plot_FDReg_hist(pval,res.qvalue$pi0,type=2,title="p values",xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"p values",line=1)
altnullhist(pval,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=FALSE,tck=-0.01)
altnullhist(pval,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8)
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"p values",line=2)

#z score histograms
nullalthist(zscore,qval_fdr,main="z scores: qvalue",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
#mtext(side=3,"z scores",line=1)
nullalthist(zscore,res.locfdr$fdr,main="locfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,res.mixfdr$fdr,main="mixfdr",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=FALSE,tck=-0.01)
nullalthist(zscore,get_lfdr(res.ash),main="ash",ncz=ncz,xaxt='n',cex.axis=0.8,ylim=c(0,0.3),xlim=c(-6,6))
axis(side=1, labels=TRUE,tck=-0.01,cex.axis=0.8)
#mtext(side=1,"z scores",line=2)

dev.off()
```

>>>>

```{r chunk-options, include=FALSE}
library(ashr)
library(qvalue)
library(locfdr)
library(mixfdr)
source("chunk-options.R")
source("../R/plot_FDReg_hist.R")
source("../R/nullalthist.R")
```

@@

Generate a small data set, and run the different methods on this data set.

@@

# Roughly compute a local fdr for qvalue (to aid the plotting of the
# decomposition in zscore space) in each bin of histogram. I set the 
# threshold at 1 so that fdr is never larger than 1.
temp     = hist(pval,nclass = 50)
bin_fdr  = res.qvalue$pi0/temp$density 
qval_fdr = bin_fdr[as.numeric(cut(pval,temp$breaks))]
qval_fdr = pmin(1,qval_fdr)
qval_fdr = qvalue::lfdr(pval)
```

Here's an example of how qvalue decomposes $p$ values into null and alternative components. First the $p$ values:

```{r plot_pvalues}
plot_FDReg_hist(pval,res.qvalue$pi0,type = 1,title = "p values",
                cex.axis = 0.8,cex.main = 0.8)
```

Now plot the decomposition:

@@

Show the distribution of the p-values and the z-scores, stratified by their
classification into the "null" and "alternative".

@@

# p value histograms
altnullhist(pval,qval_fdr,main = "p values",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = FALSE,tck = -0.01)
altnullhist(pval,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',cex.axis = 0.8)
axis(side = 1, labels = TRUE,tck = -0.01,cex.axis = 0.8)

# z score histograms
nullalthist(zscore,qval_fdr,main = "z scores",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.locfdr$fdr,main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.mixfdr$fdr,main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,get_lfdr(res.ash),main = "",ncz = 50,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
```

@@

```{r decomp_ZA_poster, dev='pdf', fig.width=6, fig.height=2, crop=TRUE}
# pdf("../figures/decomp_ZA_poster.pdf",width=6,height=2)
layout(matrix(1:8,ncol = 4,byrow = TRUE))
par(mar = c(1.3,2,1.5,0.2),mgp = c(3, 0.5, 0))
ncz = 25

# p value histograms
altnullhist(pval,qval_fdr,main = "p values: qvalue",ncz = ncz,
            xaxt = 'n',cex.axis = 0.8)
# plot_FDReg_hist(pval,res.qvalue$pi0,type = 2,title = "p values", 
#                 xaxt = 'n',cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
# mtext(side = 3,"p values",line = 1)
altnullhist(pval,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
altnullhist(pval,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = FALSE,tck = -0.01)
altnullhist(pval,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8)
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
# mtext(side = 1,"p values",line = 2)

#z score histograms
nullalthist(zscore,qval_fdr,main = "z scores: qvalue",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
# mtext(side = 3,"z scores",line = 1)
nullalthist(zscore,res.locfdr$fdr,main = "locfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,res.mixfdr$fdr,main = "mixfdr",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = FALSE,tck = -0.01)
nullalthist(zscore,get_lfdr(res.ash),main = "ash",ncz = ncz,xaxt = 'n',
            cex.axis = 0.8,ylim = c(0,0.3),xlim = c(-6,6))
axis(side = 1,labels = TRUE,tck = -0.01,cex.axis = 0.8)
# mtext(side = 1,"z scores",line = 2)
# dev.off()
```
